### 2020-05-22 ms/suppmat.tex
Add Supplementary Materials

\documentclass[12pt]{article}
\usepackage[top=0.85in,left=1.0in,right=1.0in,footskip=0.75in]{geometry}
%\usepackage[parfill]{parskip}
\usepackage{setspace}
\usepackage{lineno}
\usepackage[hidelinks]{hyperref}
%\onehalfspacing
\doublespacing

\usepackage[round,sectionbib]{natbib}
\setcitestyle{authoryear}
\bibpunct{(}{)}{;}{a}{}{;}
\bibliographystyle{fishfishnourl}

%\usepackage{textcomp}
%\usepackage{libertine}
%\usepackage{inconsolata} % sans serif typewriter

\usepackage{mathtools} % for dcases
\usepackage{xcolor} % for textcolor
\usepackage{makecell} % for \makecell and within cell line breaks (\thead)

\usepackage[T1]{fontenc}

%\makeatletter\let\expandableinput\@@input\makeatother % expandable input for \input inside tables

% Linux Libertine:
\usepackage{textcomp}
\usepackage[sb]{libertine}
\usepackage[varqu,varl]{inconsolata}% sans serif typewriter
\usepackage[libertine,bigdelims,vvarbb]{newtxmath} % bb from STIX
\usepackage[vvarbb]{newtxmath} % bb from STIX, removed bigdelims for ScholarOne rendering
\usepackage[cal=boondoxo]{mathalfa} % mathcal
%\useosf % osf for text, not math
%\usepackage[supstfm=libertinesups,%
%  supscaled=1.2,%
%  raised=-.13em]{superiors}

\usepackage{xspace}
\usepackage{xfrac} % for diagonal inline fractions in text
%\usepackage{array} % for making whole row bold in table
\usepackage{colortbl} % for background colours in table rows
\usepackage{longtable}
\usepackage{amssymb} % for \checkmark 
\usepackage{amsmath} % for \checkmark 
\usepackage{rotating}
\usepackage[nolists,tablesfirst]{endfloat} % for putting figs and tables at end of document
\DeclareDelayedFloatFlavor{sidewaystable}{table}
\usepackage{makecell} % for \makecell in tables
\usepackage{doi}

%\usepackage{xr} % to obtain label references from supp materials file
%\externaldocument[S-]{suppmat}

\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{array} % for table wrapping of columns
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
%\usepackage[detect-all]{siunitx} % for SI units

% custom hyphenation:
\hyphenation{inverse}
\hyphenation{At-lantic}
\hyphenation{elasmo-branchs}

% Macros
\newcommand{\So}{$S_{1}$\xspace}
\newcommand{\St}{$S_{2}$\xspace}
\newcommand{\Pg}{$P_r$\xspace}
\newcommand{\prmu}{$\mu_r$\xspace}
\newcommand{\prsig}{$\sigma_r$\xspace}
\newcommand{\Linf}{$L_{\infty}$}
\newcommand{\DWinf}{$DW_{\infty}$}
\newcommand{\alphat}{$\tilde{\alpha}$}
\newcommand{\lamat}{$l_{\alpha_{mat}}$}
\newcommand{\lamatb}{$l_{\alpha_{mat}}b$}
\newcommand{\rmax}{$r_{max}$\xspace}
\newcommand{\ageratio}{$\alpha_{mat}/\alpha_{max}$}
\newcommand{\yr}{year\textsuperscript{-1}}
\newcommand{\rsq}{$R^2$\xspace}
\newcommand{\mytilde}{\raise.17ex\hbox{$\scriptstyle\mathtt{\sim}$}}
% Select what to do with command \comment:  
%\newcommand{\comment}[1]{}  % comment not shown
\newcommand{\comment}[1]{\par {\bfseries \color{blue} #1 \par}} % comment shown
%% END MACROS SECTION

@@

\section*{Trends in marine survival of Atlantic salmon in eastern Canada}

\textbf{Sebasti\'{a}n A. Pardo\textsuperscript{1*}, 
        Geir H. Bolstad\textsuperscript{2}, 
        J. Brian Dempson\textsuperscript{3}, 
        Julien April\textsuperscript{4}, 
        Ross A. Jones\textsuperscript{5}, 
        Martha J. Robertson\textsuperscript{3}, 
        Dustin Raab\textsuperscript{6}, 
Jeffrey A. Hutchings\textsuperscript{1}}

\noindent\small{\textsuperscript{1} Department of Biology, Dalhousie University, Halifax, NS, Canada\\}
\small{\textsuperscript{2} Norwegian Institute for Nature Research (NINA), Trondheim, Norway\\}
\small{\textsuperscript{3} Fisheries and Oceans Canada, St. John's, NL, Canada\\}
\small{\textsuperscript{4} Minist\`{e}re des For\^{e}ts, de la Faune et des Parcs, Qu\'{e}bec, QC, Canada\\}
\small{\textsuperscript{5} Fisheries and Oceans Canada, Moncton, NB, Canada\\}
\small{\textsuperscript{6} Fisheries and Oceans Canada, Dartmouth, NS, Canada\\}
\small{\textsuperscript{*} Corresponding author: spardo@dal.ca}

\section*{Abstract}

Declines in wild Atlantic salmon (\emph{Salmo salar}) abundance throughout the north
Atlantic are primarily attributed to declining survival at sea. This
hypothesis has proven challenging to test on a river-by-river basis because of
the need to model data on both migrating smolts and returning adults and to
simultaneously estimate multiple parameters, especially for salmon spending
more than one winter at sea (1SW) before spawning. We fit a hierarchical Bayesian
maturity schedule model to data (19-42 years) for seven populations in  
Newfoundland and Labrador (NL), Qu\'{e}bec (QC), New Brunswick (NB), and Nova
Scotia (NS), Canada. We estimate survival in the first (\So) and second year at sea (\St),
and the proportion returning as 1SW adults (\Pg). Trends were not consistent
among rivers. Since 1990, \So increased at Western Arm Brook (NL)
and Rivi\`{e}re Saint-Jean (QC), but declined at Conne River (NL) and Rivi\'{e}re
de la Trinit\'{e} (QC). Since the mid-1990s, \So increased at Campbellton River (NL),
declined at LaHave River (NS), and fluctuated at Nashwaak
River (NB). Estimates of \St were highly uncertain, particularly for
1SW-dominated populations; \Pg was generally stable. These results challenge the
narrative that marine survival has changed in a temporally consistent manner
among spatially disparate populations. Our findings suggest that, at the
population level, changes in abundance are attributable to temporal shifts in
multiple components of individual fitness, including at-sea survival. If
salmon populations do not respond in a consistent, uniform manner to changing
environmental conditions throughout their range, future research initiatives
should explore why.

% The marine phase of anadromous Atlantic salmon (\emph{Salmo salar}) is the
% least known yet one of the most crucial with regards to population
% persistence. Declines in many Atlantic salmon populations in eastern
% Canada have often been attributed to changes in conditions at sea, negatively
% affecting their survival. However, marine survival estimates are difficult to
% obtain given that many individuals spend multiple winters in the ocean before
% returning to freshwater to spawn, necessitating the estimation of multiple
% parameters. To do so, we fit a hierarchical Bayesian maturity schedule model
% to smolt and adult abundance time series for seven populations located in
% Newfoundland and Labrador (NL), Qu\'{e}bec (QC), New Brunswick (NB), and Nova
% Scotia (NS). The datasets ranged between 19 and 42 years in length. We
% estimated three components of marine survival: survival in the first year at
% sea (\So), survival in the second year at sea (\St), and proportion returning
% as one sea-winter adults (\Pg). Controlling for time frame, trends in
% estimates of \So were not consistent among rivers.
% In the four populations for which data extended to 1990, marine survival
% during the first year at sea predominantly increased over time in Western Arm
% Brook (NL) and Rivi\`{e}re Saint-Jean (QC) but largely declined in Conne River
% (NL) and Rivi\`{e}re Trinit\'{e} (QC). In the three other populations, \So
% exhibited an increase in Campbellton River (NL), a decline in LaHave River
% (NS), and fluctuating stability in Nashwaak River (NB) since approximately the
% mid-1990s. 
% Estimates of \St were highly uncertain, particularly for 1SW-dominated rivers
% (Conne, Campbellton and WAB, where the abundance of 2SW returns is very low)
% and thus the posterior distributions matched our choice of prior and trends
% could not be assessed. 
% \Pg was temporally stable within rivers, except for LaHave and Nashwaak Rivers. 
% Our findings challenge the assumption that marine survival has changed in a consistent
% manner among populations across a broad geographic scale, and suggest that
% trends can be river-specific. 
% The well established correlations between climate variables and abundance are
% not being mediated solely by marine survival, and there can potentially be
% some important indirect effects on fecundity.
% Further work exploring potential correlates of marine survival and the
% potential non-lethal effects of climate effects is warranted.

Keywords: salmonid, survival at sea, natural mortality, marine mortality

%Running Head:

\section*{Introduction} % (4-5 paragraphs)

%1. Declining salmon numbers

Reductions in fishing mortality, albeit necessary, are not always sufficient
to facilitate population recovery. Experience with numerous commercially
exploited marine fisheries since the early 1990s has shown that not all
populations respond as favourably as anticipated to major reductions in
exploitation \citep{Hutchings2017}. Gradual efforts to close commercial
Atlantic salmon (\emph{Salmo salar}) fisheries in eastern Canada culminated in
full moratoria in all regions, beginning in the Maritime provinces (1984) and
following in Newfoundland (1992), Labrador (1998), and Qu\'{e}bec (2000). Since
these closures, many populations have not increased as 
expected \citep{Dempson2004, ICES2019}; some 
have been assessed as threatened or endangered by the 
Committee on the Status of Endangered Wildlife in Canada \citep[][]{Cosewic2010}, 
Canada's science advisory body (to the national government) on
species risk of extinction.
While it is not fully understood what is driving population declines, the potential
drivers of these are many \citep[see ][for a detailed discussion of possible
causes]{Cairns2001}, including but not limited to: fishing mortality \citep{Dempson2004}, 
damming of waterways and changes in the freshwater habitat \citep{Dunfield1985,Clarke2014a}, 
acidification \citep[particularly in Nova Scotia's Southern Uplands, see][]{Gibson2010}, 
predation by seals and birds \citep{Cairns2000}, negative
effects of interbreeding or interactions with escaped farmed salmon
\citep{Keyser2018}, and climate-driven changes in survival and productivity \citep{Mills2013}.

% 2.
Over the past three decades, a narrative has emerged that marine survival of
Atlantic salmon has declined throughout the North Atlantic \citep{ICES2019}.
%\citep{Hansen1998,OMaoileidigh2003,Chaput2012a}.
Based on multiple lines of evidence that climate conditions can directly and
indirectly influence the abundance and productivity of Atlantic salmon
populations \citep{Mills2013,Almodovar2019}, it has been presumed that oceanic climate effects are
driving population dynamics primarily through changes in marine survival.
Thus, an implicit assumption is that populations sharing a particular oceanic
route during seaward migration will also show similar trends in marine
survival. 
Put another way, given that salmon from different rivers 
are hypothesized to share marine habitat during some of their time at sea, it
has been presumed that populations share similar temporal trends in
at-sea mortality \citep{Friedland1993, Friedland1998, Russell2012}.  
A recent study by \citet{Olmos2019} suggested that trends in post-smolt
survival, when estimated at the stock level, are synchronously declining
for all Atlantic salmon in eastern North America.

% This perception is widespread both in the scientific
% literature as well as federal reports. The latest status report on Atlantic
% salmon by COSEWIC states that ``While the mechanism(s) of marine mortality is
% uncertain, what is clear is that the recent period of poor sea survival is
% occurring in parallel with many widespread changes in the North Atlantic
% ecosystem.'' \citep{Cosewic2010}. Consequently, there have several lines of
% inquiry as to what the causes behind these declines might be attributable to
% \citep{Friedland1993, Friedland1998}.

In contrast to the narrative of widespread, demographically similar increases
in at-sea mortality, the conservation status of Canadian salmon populations differs
considerably. Populations in the southern part of their range are more
likely to be assessed as being of conservation concern than those in more
northerly regions \citep{Cosewic2010}. This latitudinal disparity in status
suggests that if marine survival has been, or is, a key factor responsible for
most population declines, these changes are not uniformly distributed across
all populations.

% Nonetheless, most of the populations declines assessed by COSEWIC have
% occurred in populations in the southern extent of its distribution:
% populations in the Bay of Fundy, Anticosti Island, and the Atlantic coast of
% Nova Scotia being assessed as Endangered, populations in the south coast of
% Newfoundland assessed as Threatened, And the populations in the New Brunswick
% and Qu\`{e}bec coasts of the Gulf of St. Lawrence assessed as Special Concern
% \citep{Cosewic2010}. On the other hand, northermost populations have shown
% stable, or even increasing population trends, suggesting that, if marine
% survival were to be a factor in many of these declines, these changes are not
% uniformly distributed across all populations.

% While the purported decline of marine survival is mentioned widely in the 
% scientific literature, only relatively few studies have quantified it in detail.
% The basis for this premise is a limited number of time series that exhibit
% temporal declines in a proxy of marine survival (i.e. return rates or
% post-smolt survival).

% 3. Nonetheless, this premise has never been examined in detail
Given the logistical challenges associated with estimating at-sea survival, it
is not surprising that the number of studies that have estimated temporal
trends has been limited. An additional limitation has been the derivation of
proxies (e.g., return rates), rather than direct model-based estimates, of
marine survival.
\citet{Chaput2012a}, for example, examined the return rate of smolts to adult salmon 
as a metric of marine survival, finding that most Canadian populations 
had experienced declining return rates. 
However, return rates cannot directly be interpreted as marine survival rates and 
examination of trends in return rates alone can mask changes in
differential survival during different years at sea, as well as changes in the
proportion of adults returning after one or two years at sea \citep{Hubley2011}.
Recently \citet{Olmos2019} suggested that trends in post-smolt
survival, when estimated at the stock unit level, are synchronously declining
for all Atlantic salmon in Eastern North America and Canada, a conclusion ultimately 
grounded on the veracity of highly variable stock-recruitment relationships.

%4. 
In the present study, we compile data on the number of migrating smolts and number of returning adults 
for seven wild Canadian populations of Atlantic salmon to model trends in marine survival.
While some studies have previously used maturity-schedule models to estimate marine
survival for a limited number of salmon populations \citep{Chaput2003b,Hubley2011}, none
have incorporated data extending over multiple decades, nor have they examined
trends among more than two or three populations. 
Here, we develop a hierarchical Bayesian model that uses Murphy's maturity
schedule method \citep{Murphy1952}, in conjunction with informative priors, to estimate yearly
marine survival in salmon. In addition to accounting for observation error in
smolt and return estimates, we estimate the proportion of salmon returning
after one winter hierarchically.

\section*{Methods}

\subsection*{Data}

We obtained time series data of outmigrating smolt and returning adult
abundances for seven Atlantic salmon populations in eastern Canada, encompassing a
wide range of the species' distribution (Fig.~\ref{fig:map}). 
Populations included the LaHave River in Nova Scotia's (NS) Southern Uplands, 
Nashwaak River, New Brunswick (NB), Rivi\`{e}re de la Trinit\'{e} (Trinit\'{e}) and
Saint-Jean rivers in Qu\'{e}bec (QC), and  Western Arm Brook (also referred to as  WAB), 
Campbellton, and Conne River, Newfoundland (NL). 
Data were collected in NS, NB, and Newfoundland
and Labrador (NL) by Fisheries and Oceans Canada (DFO) and in QC
by the Minist\`{e}re des For\^{e}ts, de la Faune et des Parcs, Qu\'{e}bec.

\begin{figure}[htbp] \centering
    \includegraphics[width=0.85\linewidth]{figures/rivers-map2.png}
    \caption{Locations of the seven rivers in eastern Canada with time series abundance data of outmigrating smolts and 
    returning adults.} \label{fig:map} 
\end{figure}

\subsubsection*{Smolt and adult return abundance data}

Smolt and adult return abundance estimates originate from a variety of
sources. Smolt estimates from the Trinit\'{e}, Saint-Jean, LaHave, Nashwaak,
and Conne populations were obtained using a mark-recapture approach, while
estimates from the WAB and Campbellton populations were obtained by direct
counts using fish counting fences.
For further details on the data collection methodologies refer to 
\citet{Dempson1991, Venoitt2018} for NL populations, 
\citet{April2018}  for QC populations,
\citet{Jones2014} for NB population,
and \citet{Gibson2009} for NS population.

Annual return data are often recorded in terms of two size groups: small ($< 63$ cm
FL) and large ($\geq 63$ cm FL) salmon, as these closely represent different
life-history strategies (i.e. 1SW and 2SW), but can be confounded with repeat
spawners of different sizes. To correct for this in returns 
reported as small and large salmon, rather than 1SW and 2SW fish, we estimated the abundance
of 1SW and 2SW returns using yearly scale age data of a subsample of returns:

\begin{equation}
    p_{r,t,a} = \frac{\sum_{s}{(\frac{n_{r,t,s,a}}{n_{r,t,s}} * N_{r,y,s})}}{\sum_{s}{N_{r,t,s}}}
\end{equation}

where $p_{r,t,a}$ is the proportion of annual returns in river $r$, year $t$,
and of spawning history $a$ (either 1SW or 2SW returns); $n_{r,t,s,a}$ is the
number of samples in river $r$, year $t$, of returning age $a$, and of size
group $s$; $n_{r,t,s}$ is the total number of samples in river $r$, year $t$,
and of size group $s$; and $N_{r,t,s}$ is the returns of salmon
in river $r$, year $t$, and of size group $s$.
While, ideally, we would account the uncertainty in the conversion between
size group and returning age, this is being ignored by doing this conversion
deterministically as shown above. However, sampling a parameter from a
hypergeometric distribution is not possible in programming the model was
written in (i.e. Stan).

\subsection*{Bayesian model}

We developed a hierarchical Bayesian model that uses Murphy's maturity
schedule method, in conjunction with informative priors, to estimate annual
marine survival in seven populations of Atlantic salmon. We account for
observation error in smolt and return estimates, as well as estimating the
proportion returning after one winter (i.e. \Pg) hierarchically.
There is an identifiability problem in the maturity schedule equations where
the parameter estimates cannot be optimally solved \citep{Chaput2003a}.
However, this issue can be mathematically overcome, at least partially, by
using informative priors for the two marine survival parameters and the maturation
parameter in a Bayesian framework.
This method requires abundance estimates of smolts as well as abundance estimates
of returning one-sea-winter (1SW) and two-sea-winter adults. With these data,
it estimates three parameters: survival in the first year at sea (\So), survival
in the second year at sea (\St), and the proportion of fish returning after one
year at sea (\Pg).

Our model does not include repeat spawners and assumes that no fish spend
three or more winters at sea before returning to spawn for the first time.
The model also assumes that mortality in the second winter at sea (\St) for 2SW returns
is additive to mortality in the first winter at sea in the previous year, 
and therefore does not account for differences in environmental conditions experienced
between 1SW and 2SW fish of the same smolt cohort during their overlapping first year at sea.
In other words, our model assumes that the decision of returning occurs just before
actually being counted as returns and that \St is any additional mortality in
the subsequent year.

%\comment{Other assumptions...}

Observed smolt estimates were modelled hierarchically and included
observation error:

\begin{equation}
log(smolts_{obs,t,r}) = log(smolts_{true,t,r}) + \epsilon_{t,r}
\end{equation}

where $smolts_{true,t,r}$ are the true smolt abundances for year $t$ and river
$r$, and $\epsilon_{t,r}$ is the error term, which is empirically derived from
by calculating the yearly coefficient of variation in the empirically derived
smolt estimates (see Table S1 in the Supplementary material). 
Where available, we used population-specific measurement error estimates for smolt abundances; if not 
available, we set measurement error at 5\%. 
The log-transformed true smolt abundances are
normally distributed around a population-level mean and standard deviation:

\begin{equation}
log(smolts_{true,t,r}) \sim Normal(\mu_{smolts,r}, \sigma_{smolts,r})
\end{equation}

where $\mu_{smolts,r}$ and $\sigma_{smolts,r}$ are parameters estimated by the
model for each river.

Once we have yearly estimates of smolt, 1SW, and 2SW abundances, we estimate
marine survival parameters using Murphy's maturity schedule method
\citep{Murphy1952, Ricker1975}:

\begin{align}
    R_{1,t} &= smolts_{true,t-1} * S_{1,t} * Pr_t \label{eq:1}, \\
    R_{2,t+1} &= smolts_{true,t-1} * S_{1,t} * (1 - Pr_t) * S_{2,t+1} \label{eq:2}
\end{align}

where $R_{1,t}$ and $R_{2,t+1}$ are the estimated abundances of 1SW and 2SW
salmon returning in years $t$ and $t+1$, respectively, $smolts_{true,t-1}$ is the
estimated number of outmigrating smolts in year $t-1$, $S_{1,t}$ is the proportion of
salmon surviving in their first year ($t$) at sea, $Pr_t$ is the proportion of
salmon that return to spawn at year $t$, $S_{2,t+1}$ is the survival in their
second year at sea of the same cohort of salmon who did not return to spawn at
year $t$.

However, to allow for normally and log-normally-distributed parameters we
log-transform equations~\ref{eq:1} and~\ref{eq:2} to obtain

\begin{align}
    log(R_{1,r,t}) &= log(smolts_{r,t-1}) + log(Pr_{r,t}) - Z_{1,r,t} \label{eq:3}, \\
    log(R_{2,r,t+1}) &= log(smolts_{r,t-1}) - Z_{1,r,t} + log(1 - Pr_{r,t})  - Z_{2,r,t+1} \label{eq:4} 
\end{align}

where $Z_{1,r,t}$ and $Z_{2,r,t+1}$ are the instantaneous mortality rates and
$R_{1,r,t}$ and $R_{2,r,t+1}$ are the estimated 1SW and 2SW returns in
consecutive years, respectively. Observation error was included as the
standard deviation of the log-transformed return estimates from
equation~\ref{eq:4}:

\begin{align}
log(R_{obs,1,t,r}) &\sim Normal(log(R_{1,t,r}), \epsilon_{1,r}), \\
log(R_{obs,2,t,r}) &\sim Normal(log(R_{2,t,r}), \epsilon_{2,r}) \label{eq:5} 
\end{align}

where $R_{obs,1,t,r}$ and $R_{obs,2,t,r}$ are the observed return estimates
for year $t$ and river $r$ of 1SW and 2SW fish, respectively, $\epsilon_{1,r}$
and $\epsilon_{2,r}$ are the process error terms. 
These error terms are normally distributed with a standard deviation
of 0.01, which are almost equivalent to using a coefficient of variation in
return estimates 1\%, given that return estimates in equations~\ref{eq:3}
and~\ref{eq:4} are log-transformed:

\begin{align}
\epsilon_{1,r} &\sim Normal(0, 0.01) \\
\epsilon_{2,r} &\sim Normal(0, 0.01).
\end{align}

Furthermore, we use instantaneous mortality rates in the model instead of
survival probabilities as the model is more efficient in its parameter search
in log-space. Instantaneous rates are easily converted to survival
probabilities by

\begin{align}
 S_{1} &= e^{-Z_1}, \\
 S_{2} &= e^{-Z_2}. 
\end{align}

The priors for $Z_1$ and $Z_2$ are specified as log-normal distributions:

\begin{align}
Z_1 &\sim logNormal(1, 0.22),   \\ 
Z_2 &\sim logNormal(0.2, 0.3).
\end{align}

These priors, when converted to yearly survival, are roughly constrained between
0 and 0.2 for \So and between 0.2 and 0.5 for \St (Fig.~\ref{fig:priors}a).

We estimate population-level mean logit \Pg values around which the yearly \Pg
values are normally distributed. We specify different informative hyperpriors
for \prmu and \prsig based on whether the population is 1SW-dominated (i.e.
with a proportion of 1SW fish in the total returns greater than 0.9) or not:

\begin{align}
    logit(P_{r,t}) &\sim Normal(\mu_r, \sigma_r) \\
    \mu_r &\sim 
    \begin{cases}
       Normal(2.3, 0.4),  &\text{for 1SW-dominated populations} \\
       Normal(0, 2.8), &\text{for non-1SW-dominated populations} \\
   \end{cases} \\
    \sigma_r &\sim halfNormal(0, 1).
\end{align}

These priors, when converted to proportion returning, are narrowly constrained
for 1SW-dominated populations, but relatively wide for all other populations
(Fig.~\ref{fig:priors}b).

\begin{figure}[htbp] \centering
    \includegraphics[width=0.55\linewidth]{figures/priors.png} \caption{Priors
        for a) marine survival of one sea-winter (\So) and two sea-winter
        (\St) returns, and b) probability of returning as 1SW for
        1SW-dominated rivers (proportion 1SW > 0.9) and other rivers.}
    \label{fig:priors} 
\end{figure}

The model was written in Stan \citep{Carpenter2017} and run in R version 3.6.1
\citep{RCoreTeam2019} using the \texttt{rstan} package version 2.19.2
\citep{StanDevelopmentTeam2019}.
The model was run with three chains and 3,000 iterations, with the first 1,500
discarded as a burn-in. The models were considered to have converged when the
$\hat R$ of all parameters were lower than 1.03 and the effective sample size 
were higher than 500.

\subsection*{Correlations among trends in survival and parameters}

%\comment{How do were quantitatively compare trends among rivers? DFA is an
%option, but so is a t-test of averages before and after a certain year}

We looked at the correlation among trends in \So between rivers by directly estimating the 
correlation for each posterior iteration and then calculating the distribution of correlation values:

\begin{equation}
cor(Z_{1,r=1},Z_{1,r=2}) = \frac{cov(Z_{1,r=1}, Z_{1,r=2})}{\sqrt{var(Z_{1,r=1})}*\sqrt{var(Z_{1,r=2})}}\label{eq:corz1}
\end{equation}

where $Z_{1,r=1}$ and $Z_{1,r=2}$ are the $Z_1$ estimates, across years, for a
single posterior iteration for rivers 1 and 2. 
Given that the time series do not cover the same years, and that some rivers
have missing years in the middle of the time series, only years with overlapping
$Z_1$ estimates were used for the pairwise correlation estimates.
We also estimated the correlation between the model parameters for each population using
a similar approach as in equation~\ref{eq:corz1}. However we error-corrected the correlation estimates
by including the error in the variance and covariance estimates.
To get the error corrected correlations we need the error covariance ($cov_{error}$) and error
variances ($var_{error}$). Estimates of these can be obtained by estimating the covariance of
two parameters across all values in the posterior for a given year (rather
than across years for each posterior iteration), which results in one
covariance estimate for each year in each river, which then are averaged
across years for each river. Likewise, the error variance of a parameter is estimated  
across all values in the posterior for a given year and then are averaged
across years for each river:

\begin{equation}
cor_{corrected}(P_{g,r},-Z_{1,r}) = \frac{cov(P_{g,r}, -Z_{1,r}) - cov_{error}(P_{g,r}, -Z_{1,r})}
{\sqrt{var(P_{g,r})- var_{error}(P_{g,r})}*\sqrt{var(-Z_{1,r}) - var_{error}(-Z_{1,r})}}\label{eq:corparam}
\end{equation}

Lastly, to determine what parameters best explains the number of returns in
the model (including the model uncertainty), we estimated a pseudo-$R^2$ by
calculating the squared correlation between the estimated parameters and the
estimated 1SW and 2SW returns ($R_{1,r,t}$ and $R_{2,r,t+1}$ in
equation~\ref{eq:4}).

% \begin{equation}
% cor_{error}(P_{g,r},-Z_{1,r}) = \frac{cov_{error}(P_{g,r},-Z_{1,r})}{\sqrt{var_{error}(P_{g,r})}*\sqrt{var_{error}(-Z_{1,r})}}
% \end{equation}

% However correlations are not additive so to get the error corrected correlation you have to calculate

\section*{Results}

%\subsubsection*{Model convergence}

\subsubsection*{Trends in marine survival parameters}

Trends in estimates of \So were highly variable within and among rivers
(Fig.~\ref{fig:s1-dual}). The highest median posterior estimates of \So
were for the Nashwaak River in 2006 and 2008, with values of 0.18 and 0.21,
respectively. The lowest median \So estimate was in the Trinit\'{e} in 2001,
with an estimate of \So of 0.007, while the Conne, LaHave, and Trinit\'{e} had
years where estimates of \So varied between 0.01 and 0.02 (Fig.~\ref{fig:s1-faceted}).

Trends among populations also varied: Campbellton,
Saint-Jean, and Western Arm Brook populations showed increases in \So
over time, Trinit\'{e}, Conne, and LaHave populations showed decreases,
while at Nashwaak there was an increase in median \So during the early
2000s but a decrease in the 2010s. Yearly estimates of \So had very little
variability for one sea-winter dominated populations (Conne, Campbellton, and
WAB), but were more variable (i.e. wider credible intervals) in the other
populations.

>>>>

\documentclass[12pt]{article}
\usepackage[top=0.85in,left=1.0in,right=1.0in,footskip=0.75in]{geometry}
\usepackage[parfill]{parskip}
\usepackage{setspace}
\usepackage{lineno}
\usepackage[hidelinks]{hyperref}
\onehalfspacing

\usepackage[round]{natbib}
\setcitestyle{authoryear}
\bibpunct{(}{)}{;}{a}{}{;}

% Linux Libertine:
\usepackage{textcomp}
\usepackage[sb]{libertine}
\usepackage[varqu,varl]{inconsolata}% sans serif typewriter
\usepackage[libertine,bigdelims,vvarbb]{newtxmath} % bb from STIX
\usepackage[cal=boondoxo]{mathalfa} % mathcal
\useosf % osf for text, not math
\usepackage[supstfm=libertinesups,%
  supscaled=1.2,%
  raised=-.13em]{superiors}

\usepackage[utf8]{inputenc}

\usepackage{xspace}
\usepackage{xfrac} % for diagonal inline fractions in text
%\usepackage{array} % for making whole row bold in table
\usepackage{colortbl} % for background colours in table rows
\usepackage{longtable}
\usepackage{amssymb} % for \checkmark 
\usepackage{rotating}
%\usepackage[nolists,tablesfirst]{endfloat}
%\DeclareDelayedFloatFlavor{sidewaystable}{table}
\usepackage{makecell} % for \makecell in tables

% omit bibliography (\nobibliography{...})
\usepackage{bibentry}

\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{array} % for table wrapping of columns
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\usepackage[detect-all]{siunitx} % for SI units

\makeatletter\let\expandableinput\@@input\makeatother % expandable input for \input inside tables
\setlength{\LTcapwidth}{\textwidth}

% custom hyphenation:
\hyphenation{elasmo-branchs}

% Macros
\newcommand{\So}{$S_{1}$\xspace}
\newcommand{\St}{$S_{2}$\xspace}
\newcommand{\Pg}{$P_r$\xspace}
\newcommand{\Linf}{$L_{\infty}$}
\newcommand{\DWinf}{$DW_{\infty}$}
\newcommand{\alphat}{$\tilde{\alpha}$}
\newcommand{\lamat}{$l_{\alpha_{mat}}$}
\newcommand{\lamatb}{$l_{\alpha_{mat}}b$}
\newcommand{\rmax}{$r_{max}$\xspace}
\newcommand{\ageratio}{$\alpha_{mat}/\alpha_{max}$}
\newcommand{\yr}{year\textsuperscript{-1}}
\newcommand{\rsq}{$R^2$\xspace}
% Select what to do with command \comment:  
\newcommand{\comment}[1]{}  % comment not shown
%\newcommand{\comment}[1]{\par {\bfseries \color{blue} #1 \par}} % comment shown
%% END MACROS SECTION

%\bibliographystyle{model2-names}
\bibliographystyle{fishfishnourl}

%%% MY MACROS

% Make supplementary tables and figures start with S1 and T2
% use \beginsupplement to reset count
\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

@@

\noindent
{\Large Supplementary Materials}
\newline

\noindent
{\Large for}
\newline

\noindent
{\large Trends in marine survival of Atlantic salmon populations in eastern Canada}
\newline

\noindent
Sebasti\'{a}n A. Pardo\textsuperscript{1*}, Geir H. Bolstad\textsuperscript{2}, J. Brian Dempson\textsuperscript{3}, 
        Julien April\textsuperscript{4}, Ross A. Jones\textsuperscript{5}, %Martha J. Robertson\textsuperscript{3}, 
        Dustin Raab\textsuperscript{6}, Jeffrey A. Hutchings\textsuperscript{1}
\newline

\noindent
\small{\textsuperscript{1} Department of Biology, Dalhousie University, Halifax, NS, Canada\\}
\small{\textsuperscript{2} Norwegian Institute for Nature Research (NINA), Trondheim, Norway\\}
\small{\textsuperscript{3} Fisheries and Oceans Canada, St. John's, NL, Canada\\}
\small{\textsuperscript{4} Minist\`{e}re des For\^{e}ts, de la Faune et des Parcs, Qu\'{e}bec, QC, Canada\\}
\small{\textsuperscript{5} Fisheries and Oceans Canada, Moncton, NB, Canada\\}
\small{\textsuperscript{6} Fisheries and Oceans Canada, Dartmouth, NS, Canada\\}
\small{\textsuperscript{*} Corresponding author: spardo@dal.ca}
\beginsupplement

This Supplementary Material includes additional tables to accompany the publication.

\clearpage

\subsection*{Raw smolt data used in the model}

\input{smolts_tab.tex}

\input{smoltscv_tab.tex}

\subsection*{Annual smolt abundances estimates derived from the Bayesian hierarchical model}

%\begin{table}[ht]
%\centering
%\begin{longtable}{lrrrrrr}
%\caption{Posterior annual smolt estimates for the seven populations examined. These
%                    estimates differ from the smolt data used as they include both observation error and shrinkage from
%                    the hierarchical model specification.} 
%\hline
\input{tableS1_smolts}
%\end{longtable}
%\end{table}
\clearpage

\subsection*{Population-level estimates of proportion returning after one winter at sea}

\input{tableS2_smu}

\hspace{8em}

\input{tableS3_lsmu}

\input{tableS4_lssig}

% \begin{figure}[htbp] \centering
%     \includegraphics[width=0.9\linewidth]{simul_timeseries_declining.pdf} \caption{Simulated
%         time series of returning adult salmon abundance in the six scenarios. The lines denote the
%     simulated abundance estimates without observation error while the points are the same estimates
% including observation error.}
% \label{fig:simul-ts-dec} 
% \end{figure}

% \begin{figure}[htbp] \centering
%     \includegraphics[width=0.95\linewidth]{sir_out_18_declining.pdf} \caption{Yearly
%         estimated \So, \St, and \Pg values in the six scenarios. True values
%         are denoted by blue circles, black circles show median estimates,
%         error bars indicate the 25\% and 75\% quantiles.}
% \label{fig:sir-out-dec} 
% \end{figure}

% \begin{figure}[htbp] \centering
%     \includegraphics[width=0.7\linewidth]{simul_plots_6_declining.pdf}
%     \caption{Comparison of estimated and true \So values in the six 
%         scenarios. True \So values are deterministic, black circles show
%         median \So estimates, error bars indicate the  25\% and 75\%
%         quantiles, while the blue line denotes a linear model fit of the
%         medians. The one-to-one relationship is shown by the gray dashed
%         line.} 
% \label{fig:simul-six-plots-dec} 
% \end{figure}

%\clearpage
%\subsection*{References}
\nobibliography{subset}
